Insight Into Lumbar Back Pain \subtitle{What the Lumbar Spine Tells About Your Life}

### 2014-09-09 draft.tex
Removed section about Age; Added section about BMI

no group (a), subdivision into \emph{male and female subjects} (b), quartiles of \emph{age} (c) and quartiles of \emph{body size} (d).

@@

	\item subdivision into \emph{age quantiles}, yielding the groups (21, 43.8] (43.8, 54] (54, 64] (64, 90],

@@

@@

%\com{Potentially skip these due to the confounding of body size}
\subsubsection{Age Groups}
%%
Gender could be classified for each age group using \emph{mean curvature} and \emph{curvature angle}.
%%
The accuracy varies between $0.35$ (\emph{$43.8-54$ years of age}, 3 decision rules) to $0.27$ (\emph{$21-43.8$ years of age}, 6 decision rules).
%%
The starting age of smoking could be characterized well with an error rate between $0.25$ to $0.3$ for all \emph{age} groups except of \emph{$21.1-43.8$ years of age}.
%%
The result however is probably over fitted to the data due to tree sizes between $16$ and $23$.
%%
\com{Warum?}
%%

Many variables, such as \emph{body size} can be described with an error rate of $0.35$ to $0.45$ but only using large decision trees with over 20 rules.
%Body-size can be determined with an error rate of $0,36$ to $0,39$ for subject
% age groups, but also imposes over fitting to the data yielding a decision tree of over 30 rules.
%%
Notable is also the increasing accuracy with increasing decision tree size.
%%
For subjects between \emph{$43.8-54$ years of age}, the classifier discriminated subjects with thyroid nodules with an error rate of $0.32$.

The dummy variable analysis shows many results using a decision tree with one rule based on \emph{mean curvature} with accuracy between $0.4$ and $0.5$.
%%
It shows that mean curvature can be used to predict dependencies to variables, such as \emph{high blood pressure}, \emph{hypothyroidism} or \emph{nutrition}.

@@

\section{\uppercase{Conclusion}}
\label{sec:Conclusion}
\noindent The presented results indicate that the parameters \emph{torsion}, \emph{curvature} and \emph{curvature angle} are not enough to characterize pathological changes in the lumbar spine.
%%
They are well suited to characterize subject gender due to their body size difference using \emph{mean curvature} and \emph{mean curvature angle}.
%%
This is due to the higher curvature of smaller subjects and the significant difference of average \emph{body size} between \emph{males} and \emph{females}.
%%
Gender was characterized well in the analysis of all subjects as well as subjects divided by \emph{age}.

>>>>

%no group (a), subdivision into \emph{male and female subjects} (b), quartiles of \emph{age} (c) and quartiles of \emph{body size} (d).
no group (a), subdivision into \emph{male and female subjects} (b), quartiles of \emph{Body-Mass-Index} (\emph{BMI}) (c) and quartiles of \emph{body size} (d).

@@

	\item subdivision into \emph{Body-Mass-Index quantiles} ($BMI = \frac{m}{l^2}$ where $m$ is the \emph{body mass} in kilogram and $l$ is the \emph{body size} in meter), yielding the groups (17, 24.7] (24.7, 27.4] (27.4, 30.5] (30.5, 48],

@@

%The \emph{BMI} is calculated using the formula $BMI = \frac{m}{l^2}$ where $m$ is the \emph{body mass} in kilogram and $l$ is the \emph{body size} in meter.

@@

%\com{Potentially skip these due to the confounding of body size}
\subsubsection{Body-Mass-Index Groups}
%%
Gender could be classified for each \emph{BMI} group using \emph{mean curvature} and \emph{curvature angle}.
%%
The accuracy varies between $0.31$ (\emph{BMI} of \emph{$30.5-48$}, 4 decision rules) to $0.39$ (\emph{BMI} of \emph{$24.7-27.4$}, 5 decision rules).
%%
The starting age of smoking could be characterized well with an error rate between $0.25$ to $0.32$ for all \emph{BMI} groups except of subjects with BMI of \emph{$30.5-48$}.
%%
The result however is probably over fitted to the data due to tree sizes between $14$ and $16$.
%%
\com{Warum?}
%%

Some variables, such as \emph{body size} can be described with an error rate of $0.3$ to $0.36$ but only using large decision trees with over 20 rules.
%%
Notable is also the increasing accuracy with increasing decision tree size.
%%
Using mostly \emph{mean curvature} and \emph{curvature angle}, the leg pain level can be predicted using $14$ rules with an error rate of $0.46$ for obese subjects (\emph{BMI} higher than $30$).
%%
The dichotomous variable whether the subject has felt pain in the last seven days can also be predicted for this group using the same features.
%%
The resulting tree consists of $8$ rules and has an error of $0.35$.
%%
Obese subjects are prone to back and leg pain due to a more stressed lumbar spine.
%%
How much the spine is deformed under the heavy stress seems to directly influence the pain levels for these subjects.

The dummy variable analysis shows many results using a decision tree with one rule based on \emph{mean curvature} or \emph{curvature angle} with accuracy between $0.35$ and $0.47$.
%%
%It shows that mean curvature can be used to predict dependencies to variables, such as \emph{high blood pressure}, \emph{hypothyroidism} or \emph{nutrition}.

% \subsubsection{Age Groups}
% %%
% Gender could be classified for each age group using \emph{mean curvature} and \emph{curvature angle}.
% %%
% The accuracy varies between $0.35$ (\emph{$43.8-54$ years of age}, 3 decision rules) to $0.27$ (\emph{$21-43.8$ years of age}, 6 decision rules).
% %%
% The starting age of smoking could be characterized well with an error rate between $0.25$ to $0.3$ for all \emph{age} groups except of \emph{$21.1-43.8$ years of age}.
% %%
% The result however is probably over fitted to the data due to tree sizes between $16$ and $23$.
% %%
% \com{Warum?}
% %%
%
% Many variables, such as \emph{body size} can be described with an error rate of $0.35$ to $0.45$ but only using large decision trees with over 20 rules.
% %%
% Notable is also the increasing accuracy with increasing decision tree size.
% %%
% For subjects between \emph{$43.8-54$ years of age}, the classifier discriminated subjects with thyroid nodules with an error rate of $0.32$.
%
% The dummy variable analysis shows many results using a decision tree with one rule based on \emph{mean curvature} with accuracy between $0.4$ and $0.5$.
% %%
% It shows that mean curvature can be used to predict dependencies to variables, such as \emph{high blood pressure}, \emph{hypothyroidism} or \emph{nutrition}.

@@

\section{\uppercase{Conclusion}}
\label{sec:Conclusion}
\noindent The presented results indicate that the parameters \emph{torsion}, \emph{curvature} and \emph{curvature angle} are not enough to characterize pathological changes in the lumbar spine.
%%
They are well suited to characterize subject gender due to their body size difference using \emph{mean curvature} and \emph{mean curvature angle}.
%%
This is due to the higher curvature of smaller subjects and the significant difference of average \emph{body size} between \emph{males} and \emph{females}.
%%
Gender was characterized well in the analysis of all subjects as well as subjects divided by \emph{BMI}.
